Percussive Sound Generation with Virtual Listeners and Modular Synthesizers

### 2020-08-20 templates_aimusic2020/paper.tex
Updates from Overleaf

% intro, 1 page
% background and methodology 2 pages
% implementation 
% results 2.5
\maketitle

\begin{abstract}
Digital sound artists often require a variety of percussive samples for their music. For more than two decades research involving digital synthesis, heuristic search, and neural networks has been used for the generation of novel sounds. This work is a proof of concept in automatic programming of synthesizers towards their best approximation of drum sounds. We centered our approach around the combination of two central components: \begin {enumerate*} [label=(\roman*)] \item a \emph{virtual ear} capable of evaluating the proximity of \enquote{unusual} sounds to drum sounds and \item a dynamic \emph{virtual synthesizer} with a rich set of tractable parameters\end{enumerate*}. We present a generative pipeline that utilizes classical digital signal processing functions guided by supervised learning towards the generation of one-shot percussive sounds. We present our findings and measurements of a number of approaches taken. Advantages, shortcomings and practicality of the employed methodologies are demonstrated. We share our curated dataset of sounds and codebase\footnote{\url{https://github.com/imilas/Synths_Stacks_Search}} which can be used for its further expansion.
\keywords{Automatic Synthesizer design. Machine Listening. Sound Analysis. Novelty and Originality}
\end{abstract}

@@

\subsection{Related Works}
\label{related_works}
\begin{center}
\begin{table}[h]
\resizebox{\linewidth}{!}{\begin{tabular}{|c c c c|} 
\hline
Work & Feature extraction & Synthesis & Specilization & \hline
Oord 2017 & CNN & CNN &Speech& \hline
Yee-King 2018 & LSTM on Paremters & DSP & Synth Pads & \hline
Aouameur 2019 & Latent layer& Decoding of Latent Layers & Percussion & \hline
Ramires 2020 & Latent layer & FeedForward Network & Percussion & \hline
Yamamoto 2020 & GAN & GAN & Speech & \hline
\end{tabular}}
\caption{Quick reference of recent related works}
\end{table}
\label{table:recent_works}
\end{center}
Numerous deep, neural network models have been proposed and utilized for the purpose of signal generation in recent years. WaveGans and WaveNet have been subject to significant improvements and experiments since their proposal~\cite{nsynth2017,yamamoto2020parallel,oord2017parallel}. Particularly relevant are the utilization of variational autoencoders (VAE's) for generation of percussive samples~\cite{aouameur2019neural} and generation of percussive sounds by decoding a small set of latent features~\cite{ramires2020neural}.
Automatic programming of virtual synthesizers has long been a topic of interest. In early 2000s, Interactive Genetic Algorithms (IGA's) were utilized for the generation of new sounds with various sound-engines~\cite{johnson1999exploring,dahlstedt2001creating}. More recent work by Yee-King et al.~\cite{yee2018automatic} used Long short-Term Memory (LSTM) models and genetic algorithms to find the exact parameters used to create a group of sounds. The sounds approximated were made by the same virtual synthesizer, not an external source; making the eventual replication certain even with random search. Since this work appears more focused on pads and textures rather than drums, feature matching appears to not be concerned with the envelope of the sounds but rather the frequency content within arbitrary time windows. Yet another recent work by Esling et al. used a large dataset of over 10,000 presets for a commercial VST synthesizer to learn a latent parameter space which can be sampled for creation of new audio~\cite{esling2019universal}. As stated before, our work explores the rapid approximation of percussion sounds with no previous knowledge about the sonic capabilities of our virtual synthesizer, exploring the actual parameter space rather than its latent representation.

@@

\begin{figure*}[htbp]
\label{fig_example_sine}
\centering
\includegraphics[width=0.45\linewidth,angle =-90 ]{images/periodic_function.png}
\caption{A computer can simulate waveforms by utilizing periodic functions. Digital waveforms are discrete approximations of analogue waves~\cite{mitchell2009basicsynthChap5} }
\end{figure*}

@@

\begin{figure}[htbp!]
    \begin{center}
    \textbf{Data Overview}
    \makebox[\textwidth]{
    \includegraphics[width=0.95\linewidth]{images/venn_data.pdf}}
    \end{center}
    \caption{ 
    Scales are symbolic (true scales are not measurable and/or subjective).  $\mathcal{N}$ is the set of percussive sounds a synthesizer is capable of making. Our positive samples, $\mathcal{T^{+}}$, is a small fraction of a wide variety of percussive sounds that are conceivable. The size of $\mathcal{T^{-}}$ can be set to any number(we can generate as any number of random samples), but we prefer to keep it no larger than  $\mathcal{T^{+}}$. $\mathcal{H}$ is a series of sounds sent to the ear for classification. The ideal ear will detect  $\mathcal{H^{+}}$, the overlap of $\mathcal{H}$ and $\mathcal{N}$, with high precision and recall. 
    }
\label{fig:ven_data}
\end{figure}
\hypref{Decision.1} requires knowledge of what drums \textbf{do not} sound like, or knowledge of an infinitely large set, which cannot be fully represented via examples. An important consideration is that the source of sounds used in training the model (organic drum sounds) will be fundamentally different from the source of unlabeled sounds we wish to categorize (noise from a synthesizer). This issue is reflective of the OSR problem~\cite{geng2020recent,mundt2019open}. Figure~\ref{fig:ven_data} highlights a number of caveats with this approach. we denote $\mathcal{N}$ as the set of percussive sounds a synthesizer is capable of making. $\mathcal{T^{+}}$, our positive examples, is a set of sound extracted from material drums. $\mathcal{T^{-}}$, is a small subset of an infinite set which is meant to represent all noises which the synthesizer is capable of making and likely includes sounds in $\mathcal{N}$. During the hearing test, the virtual ear continually receives synthesizer noises for classification. The set of received noises is denoted as $\mathcal{H}$. The performance of the ear is measured by its precision and recall in finding the overlap between $\mathcal{H}$ and $\mathcal{N}$. It is important to ensure that the change in learning domains ---particularly with the case of positive examples of percussive sounds--- does not interfere with transformation of knowledge from the training of the ear to the hearing test. To maximize the transference of knowledge from one domain to another, we seek to learn from agnostic feature sets that capture fundamental characteristics of the data points.
\subsection{Feature Extraction}
To train our virtual ear, we require generalizable, domain agnostic features. Various works have demonstrated effective reconstruction of signals given their Short-time Fourier Transforms (STFT) ~\cite{nawab1983signal,griffin1984signal}. If the STFT of a signal can be used its reconstruction, perhaps it can be utilized as a source of fundamental features necessary for audible signal categorization. We defined 3 STFT transformation functions which we believe to capture important, unique attributes of percussive sounds (figure ~\ref{fig:stackspectrums}). Our transformations are GPU based, allowing their application in tandem to the training of the classifier or tuning of STFT transformation during hyper-paramter optimization (see Appendix~\ref{appendix:F}).

\subsection{Ear Types}
\label{classifiers}
Two groups of classifiers are implemented: \emph{two phased ears} (TPEs) and \emph{Mixed Ear Models} (MEMs). TPEs combine different models specializing in ~\hypref{Decision.2} or ~\hypref{Decision.1} and trained only with our transformation functions. MEMs are a proof of concept attempt at addressing the OSR problem by usage of autoencoded representation of sounds as features. MEMs are trained using highly compressed, automatically encoded representation of sounds and can give simultaneous answers to both questions. Model specifications for TPE and MEMs can be found in Appendix~\ref{appendix:C}. We speculate that the features used to represent sounds plays a much more important role in determining outcomes, particularly when working with relatively small datasets.

The features used by MEMs are encodings extracted from the latent layer of an autoencoder trained with MixedDB. To better understand the delineation potential of the encoded features, 64-dimensional latent representation of 20\% of MixedDB and 1000 Noise sounds were projected onto a 2 or 3-dimensional planes using t-SNE. Referring to~\hypref{Decision.1} and~\hypref{Decision.2}, we expect the drum sounds to be clustered together, with the majority of the synthetic noise sounds appearing away from these clusters. The projections confirmed that this is the case(see Appendix~\ref{appendix:E}). We are most interested in synthetic noise sounds which appear within or near these clusters. Do these synthetic noise sounds have drum like characteristics?\\ 
To test this, we implemented interactive projection graphs. These graphs allow interactions such as playing the samples(by hovering the cursor on-top) and movement of the scene camera. Manual inspections revealed a noticeable, positive correlation between distance and similarity of a synthetic sound to drum clusters.

\section{Pipelines and Surveys}
\label{gens}
\label{surveys}
With our virtual synthesizer and virtual ears in place, we build two pipelines, one using all TPE models and the other using our best MEM model. To assess the performance of each pipeline, we randomly create drum programs and generate and the corresponding signal. This signal is then fed through the ear model to determine if the signal is percussive. If so, a category is assigned to the sound. This sound and the corresponding synthesizer program are saved to hard-disk. Here we assess the success rate of two different pipelines by manual inspection of sounds it has categorized as drums, as well as the category of those drums. The authors of this work attempt their own categorization of a randomly selected subset of each pipeline's results. We cross reference the manual categorizations with the categories assigned by the virtual ear: Do we agree that the sounds are percussive? Do we agree with the drum group assigned to the percussive sounds?

 \section{Survey of Two-Phased Ear Performance}

\begin{table}[htbp]
 \resizebox{\linewidth}{!}{\begin{tabular}{||c c c c c c c||} 
 \hline
 Drop Rule & Size & HvH & H+FC & H+CNN & H+ALL & 3 models \\ [0.5ex] 
 \hline
 No Drops & 257 &0.37 & 0.35 & 0.36 & 0.36 & 0.28\\ 
 \hline
 Assigned \enquote{Bad} By Both & 236 & 0.31 & 0.37 & 0.37 & 0.38 & 0.30 \\
 \hline
 Assigned \enquote{Bad} By Either & 180 & 0.47 & 0.50 & 0.48 & 0.48 &  0.34 \\
 \hline
 Assigned \enquote{Bad} or \enquote{Other} By Either & 154 & 0.47 & 0.59 & 0.54 & 0.50 &  0.35 \\
 \hline
\end{tabular}}
\caption{\label{kappa_table_TPE}Table of Fleiss' kappa coefficient to measure the degree of agreement between persons (HvH), persons with FC model (H+FC), persons with CNNLSTM model, persons with all models (H+ALL), and between the 3 models. \enquote{Drop Rule} column indicates if any samples were dropped. We show the measurements after dropping samples if they are deemed bad by either or both responders. We also show measurements after dropping the \enquote{other} category along with samples deemed bad by either responder. }
\end{table}
To measure the quality of the samples produced by our pipeline combined with the TP ear models, we run the model until an adequate number of samples in each category are saved to disk. Next, we randomly generated around 50 samples in the following categories: \enquote{snare}, \enquote{kick}, \enquote{hat}, \enquote{clap} and \enquote{other} (combination of rims, 
shakers and unusual percussive sounds). This gave us a total of 257 samples. These samples were determined to be percussive and then categorized by 4 different models (FC, CNNLSTM, E+F and AVG). We ensured a balanced division between samples of stack size 1, 2 and 4 (each stack is responsible for a third of the samples under each category). Both reviewers then categorized these samples without knowledge of other categorizations (person or computation models). Additionally, each responder had the option of labeling samples as \enquote{bad} for samples that they deemed not percussive.

We assess the reliability of agreement between persons and categorization models via the Fleiss' kappa coefficient \cite{fleiss1971measuring}. The value of 0 or less for this coefficient indicates no agreement beyond random chance, and the value of 1 indicating perfect agreement. Our kappa measurements showcased in~\ref{kappa_table_TPE} lie within the 0.35-0.45 range, indicating mild to moderate agreement between persons and machines. We again measure this coefficient after dropping samples that were categorized as \enquote{bad} by the authors, as samples that persons deem to be \enquote{bad} indicate a failure in Phase 1 and should not have been categorized by the models at all.

\section{Survey of Mixed-Models Ear Performance}
 Keeping the other components of the pipeline the same, we replace the two-phased ear in the previous pipeline with the MME. Our survey differs from the previous survey in that the pipeline only creates the four drum categories. Our main concern here is changes in the amount of \enquote{bad} samples. We show measurements before and after dropping all \enquote{bad} and \enquote{other} samples for a fairer comparison with the TPE pipeline.

 \begin{table}[t]
 \resizebox{\linewidth}{!}{\begin{tabular}{||c c c c||} 
 \hline
 Drop Rule & Size & HvH & H+MME \\
 \hline
 No Drop & 300 & 0.336 & 0.250\\ 
 \hline
 Assigned Bad By Both & 249 & 0.200 & 0.260 \\
 \hline
 Assigned Bad By Either & 151 & 0.460 &  0.473 \\
 \hline
Assigned \enquote{Bad} or \enquote{Other} By Either  & 120 & 0.620   &  0.587 \\
 \hline
\end{tabular}}
\caption{\label{kappa_table_MME}Table of Fleiss' kappa coefficient to measure the degree of agreement between persons (HvH) and persons and MME. We also measure the agreeability scores after dropping bad samples if both or either persons assigned the sample as such. We also measure agree-ability when all samples deemed \enquote{bad} or \enquote{other} by either person are removed.}
\end{table}

\subsection{Takeaways from These Surveys}
\label{survey2_takeaway}
The results from the two surveys are not easily comparable. However, no major improvements in agree-ability or a decrease in the number of bad samples can be seen, in fact, the opposite is more likely to be true. However, we are yielding comparable results to the TPE pipeline despite having a fraction of the features to learn from (latent layer + 10 envelope features). This shows that latent layers of an autoencoder network can be used as a low-dimensional proxy for representation of more complex inputs, despite the autoencoder not being trained for such a task. Agree-ability scores for MME pipeline are often lower, but we reiterate that this can be caused by having two out of six (rather than one out of six in the TPE pipeline) categories which are exclusive to survey responders. Supporting this hypothesis is the notable improvement in agree-ability observed after dropping all \enquote{bad} and \enquote{other} samples. Here, both persons agree with each other and with the MME at a stronger rate than in the TPE pipeline, despite having similar number of samples (120 for MME and 154 for TPE) and equal number of categories to choose from. This values are observed in the last rows of tables~\ref{kappa_table_TPE} and ~\ref{kappa_table_MME}. The higher DVD categorization accuracy achieved by the mixed-model ear compared to the best two-phaased ear is consistent with this observation: Given that the sound being categorized is a drum, the MME will yield better results that the TPE. However, the MME pipeline does not yield better results in the DVN task. 
\begin{itemize}
    \item The survey brings into question the reliability of our phase 1 models, as 30\% of the generated samples were deemed not percussive by at least 1 reviewer and 8\% by both reviewers
    \item The task of categorizing synthetic drums is difficult. Survey shows that the scale of agreement within persons as well as between persons and various model combinations is moderate at best, even after removal of \enquote{bad} samples.  While the same models can easily achieve 98+ percent accuracy when tested on sounds from physical drum samples. 
    \item While there is much room for improvement, our pipelines can generate and categorize drums and percussive sounds with a promising degree of success. 
\end{itemize}
\bibliographystyle{apacite}
\bibliography{CSMC_MUME_LaTeX_Template.bib}

\begin{appendices}
\chapter{Datasets}
\label{appendix:A}
\begin{table}[h!]
\centering
\begin{tabular}[width=\paperwidth]{|l|l|l|l|l|l|l|l|l|l|}
\hline
DB Name & kick & snare & clap & tom\_high & tom\_mid & tom\_low & hihat\_closed &  hihat\_open & rim \\ \hline
MixedDB & 648 & 732 & 118 & 179 & 139 &  188 & 187 & 280 & 105 \\\hline
\end{tabular}
\caption{Database 1: Mixed sources}
\label{db:self}
\end{table}

@@

%================
\chapter{Virtual Ear Classifier Definitions}
\label{appendix:C}
\section{Two Phased models}
Using the described features, we trained several neural network models for Phase 1 and 2 in the Pytorch environment. The task of Phase 1 is to separate drums from not-drums (DrumVsNotDrum, or DVN). The task of Phase 2 is to categorize drums and percussion (DrumVsDrum, or DVD). We kept our feature space small, making it viable for feature selection and model design to be done on a trial and error basis. For all models, accuracy is calculated by prediction of all test dataset labels and the loss function and optimizer are Categorical-CrossEntropy and Adam respectively. Training continues until no reduction in loss and accuracy is observed in 10 epochs.  All activation functions are PReLU:
\begin {enumerate}
\item FC-DVN: Fully connected network trained on Envelope features, reaching 97\% accuracy on our test data for Phase 1. With size of 10x5x10.
\item CNNLSTM-DVN: A combination of CNN and LSTM models, where the CNN model extracts higher level features that are fed temporally to an LSTM cell. This model is trained on spectrum data and reaches 98\% accuracy on our test set. Its structure is the combination of a CNN with 2 output channels and kernel size $(7,3)$; Followed by an LSTM model of hidden size 800 and a fully connected layer of size 20x2.
\item E+F-DVD: A fully connected model trained on a concatenation of envelope and frequency features. Reaching 80\% accuracy for 6-way drum categorization in Phase 2. Size of 50x10x2x6.
\item CNN-DVD: A CNN model trained on Spectrum features. Reaching 82\% accuracy in a 6-way drum categorization in Phase 2. A combination of a CNN model with output channel size of 4, kernel of size of 5, another CNN model with output channel size of 8 and kernel of size 3. Followed by a fully connected network of shape 100x20x6.
\item FC-DVD: Fully connected 3 layer neural net with 78\% accuracy for 6-way drum categorization in Phase 2. Size of 400x200x50.
\end{enumerate}
Unlike EE models, parameters are hand-picked and un-tuned. Higher accuracy rates in these models do not necessarily translate to higher agreeableness with surveyors. Since the positive sounds in the hearing test are sourced from a synthesizer, model accuracy on test data alone cannot be relied upon when the domain of sounds 
being categorized is switched.

@@

%==============
\chapter{AutoEncoder Structures and Hyper-Parameter Optimization}
% could do multi-objective pareto front stuff too
Within the context of machine learning, a model's \emph{hyper-parameters} are fixed parameters which are set before the training begins (e.g number of layers, size of layers, loss function) and are not learned during the minimization procedure~\cite{bengio2000gradient}. To assist us with the construction of our model, we defined a number of possible choices for the architecture of our model and audio-transformers and used hyper-parameter optimization to extract promising sets of values~\ref{table:hyper_params}.

@@

Using the optuna optimization tools \cite{akiba2019optuna}, we conducted 500 search trials. The trial's success is measured in their final loss value, calculated by applying the model to test data-set. Each trials consisted of 20 epoch of training using a selected set of hyper-parameters. An additional 100 trials with 40 epochs of training were conducted following the initial 500 to test the effect of longer epochs. Each trial's intermediate results (loss at every n epoch,s where $0<n<20$) were reported to a multi-armed bandit based pruner for early stoppage of unpromising trials~\cite{li2017hyperband}. We employed a tree-structured parzen estimator for better navigation of the search space~\cite{bergstra2011algorithms,akiba2019optuna} but found short reversions to a random sampling coupled with a decrease in the frequency of pruning helpful in exiting local minima. \\

>>>>

% intro, 1 page
% background and methodology 2 pages
% implementation 
% results 2.5
\maketitle
% For more than two decades research involving digital synthesis, heuristic search, and neural networks has been used for the generation of novel sounds.
\begin{abstract}
This work is a proof of concept in automatic programming of synthesizers towards their best approximation of drum sounds. Our approach is based on the combination of two central components: \begin {enumerate*} [label=(\roman*)] \item a \emph{virtual ear} capable of evaluating the proximity of \enquote{unusual} sounds to drum sounds and \item a dynamic \emph{virtual synthesizer} with a rich set of tractable parameters\end{enumerate*}. We present a modular design of a generative pipeline which utilizes (semi-)supervised learning and classical digital signal processing for the generation of one-shot percussive sounds. We present our findings and measurements of a number of approaches taken, particularly in regard to the implementation of an effective virtual ear. Advantages, shortcomings and practicality of the employed methodologies are demonstrated and discussed. We share a number of datasets, notably our curated dataset of free percussive sounds.

\keywords{Automatic Synthesizer design. Machine Listening. Sound Analysis. Novelty and Originality}
\end{abstract}

@@

\subsection{Related Works}
\label{related_works}
\begin{center}
\begin{table}[h]
\resizebox{\linewidth}{!}{\begin{tabular}{|c c c c|} 
\hline
Work & Feature extraction & Synthesis & Specilization & \hline
Oord 2017 & CNN & CNN &Speech& \hline
Yee-King 2018 & LSTM on Paremters & DSP & Synth Pads & \hline
Aouameur 2019 & Latent layer& Decoding of Latent Layers & Percussion & \hline
Ramires 2020 & Latent layer & FeedForward Network & Percussion & \hline
Yamamoto 2020 & GAN & GAN & Speech & \hline
\end{tabular}}
\caption{Quick reference of recent related works}
\end{table}
\label{table:recent_works}
\end{center}
Numerous deep, neural network models have been proposed and utilized for the purpose of signal generation in recent years. WaveGans and WaveNet have been subject to significant improvements and experiments since their proposal~\cite{nsynth2017,yamamoto2020parallel,oord2017parallel}. Particularly relevant are the utilization of variational autoencoders (VAE's) for generation of percussive samples~\cite{aouameur2019neural} and generation of percussive sounds by decoding a small set of latent features~\cite{ramires2020neural}.
Automatic programming of virtual synthesizers has long been a topic of interest. In early 2000s, Interactive Genetic Algorithms (IGA's) were utilized for the generation of new sounds with various sound-engines~\cite{johnson1999exploring,dahlstedt2001creating}. More recent work by Yee-King et al.~\cite{yee2018automatic} used Long short-Term Memory (LSTM) models and genetic algorithms to find the exact parameters used to create a group of sounds. The sounds approximated were made by the same virtual synthesizer, not an external source; making the eventual replication certain even with random search. Since this work appears more focused on pads and textures rather than drums, feature matching appears to not be concerned with the envelope of the sounds but rather the frequency content within arbitrary time windows. Yet another recent work by Esling et al. used a large dataset of over 10,000 presets for a coMEMrcial VST synthesizer to learn a latent parameter space which can be sampled for creation of new audio~\cite{esling2019universal}. As stated before, our work explores the rapid approximation of percussion sounds with no previous knowledge about the sonic capabilities of our virtual synthesizer, exploring the actual parameter space rather than its latent representation.

@@

% \begin{figure*}[htbp]
% \label{fig_example_sine}
% \centering
% \includegraphics[width=0.3\linewidth,angle =-90 ]{images/periodic_function.png}
% \caption{A computer can simulate waveforms by utilizing periodic functions. Digital waveforms are discrete approximations of analogue waves~\cite{mitchell2009basicsynthChap5} }
% \end{figure*}

@@

\begin{figure}[htbp!]
    \begin{center}
    \textbf{Data Overview}
    \makebox[\textwidth]{
    \includegraphics[width=0.95\linewidth]{images/venn_data.pdf}}
    \end{center}
    \caption{ 
    Scales are symbolic (true scales are not measurable and/or subjective).  $\mathcal{N}$ is the set of percussive sounds a synthesizer is capable of making. Our positive samples, $\mathcal{T^{+}}$, is a small fraction of a wide variety of percussive sounds that are conceivable. The size of $\mathcal{T^{-}}$ can be set to any number(we can generate as any number of random samples), but we prefer to keep it no larger than  $\mathcal{T^{+}}$. $\mathcal{H}$ is a series of sounds sent to the ear for classification. The ideal ear will detect  $\mathcal{H^{+}}$, the overlap of $\mathcal{H}$ and $\mathcal{N}$, with high precision and recall. 
    }
\label{fig:ven_data}
\end{figure}
\hypref{Decision.1} requires knowledge of what drums \textbf{do not} sound like, or knowledge of an infinitely large set, which cannot be fully represented via examples. An important consideration is that the source of sounds used in training the model (organic drum sounds) will be fundamentally different from the source of unlabeled sounds we wish to categorize (noise from a synthesizer). This issue is reflective of the OSR problem~\cite{geng2020recent,mundt2019open}. Figure~\ref{fig:ven_data} highlights a number of caveats with this approach. we denote $\mathcal{N}$ as the set of percussive sounds a synthesizer is capable of making. $\mathcal{T^{+}}$, our positive examples, is a set of sound extracted from material drums. $\mathcal{T^{-}}$, is a small subset of an infinite set which is meant to represent all noises which the synthesizer is capable of making and likely includes sounds in $\mathcal{N}$. During the hearing test, the virtual ear continually receives synthesizer noises for classification. The set of received noises is denoted as $\mathcal{H}$. The performance of the ear is measured by its precision and recall in finding the overlap between $\mathcal{H}$ and $\mathcal{N}$. It is important to ensure that the change in learning domains ---particularly with the case of positive examples of percussive sounds--- does not interfere with transformation of knowledge from the training of the ear to the hearing test. To maximize the transference of knowledge from one domain to another, we seek to learn from agnostic feature sets that capture fundamental characteristics of the data points.
\subsection{Feature Extraction}
To train our virtual ear, we require generalizable, domain agnostic features. Various works have demonstrated effective reconstruction of signals given their Short-time Fourier Transforms (STFT) ~\cite{nawab1983signal,griffin1984signal}. If the STFT of a signal can be used its reconstruction, perhaps it can be utilized as a source of fundamental features necessary for audible signal categorization. We defined 3 STFT transformation functions which we believe to capture important, unique attributes of percussive sounds (figure ~\ref{fig:stackspectrums}). Our transformations are GPU based, allowing their application in tandem to the training of the classifier or tuning of STFT transformation during hyper-paramter optimization (see Appendix~\ref{appendix:hyperparam}).
\subsection{Ear Types}
\label{classifiers}

Two groups of classifiers are implemented: \emph{two phased ears} (TPEs) and \emph{Mixed Ear Models} (MEMs). TPEs combine different models specializing in~\hypref{Decision.2} or~\hypref{Decision.1} and trained only with our transformation functions. MEMs are a proof of concept attempt at addressing the OSR problem by usage of autoencoded representation of sounds as features. MEMs are trained using highly compressed, automatically encoded representation of sounds and can give simultaneous answers to both questions. Model specifications and measurments for TPE and MEMs can be found in Appendix~\ref{appendix:C}. We speculate that the features used to represent sounds plays a much more important role in determining outcomes, particularly when working with relatively small datasets.

The features used by MEMs are encodings extracted from the latent layer of an autoencoder trained to encode and decode spectrogram representations of sound in MixedDB. To better understand the delineation potential of the encoded features, 64-dimensional latent representation of 20\% of MixedDB and 1000 Noise sounds were projected onto a 2 or 3-dimensional planes using t-SNE. Referring to~\hypref{Decision.1} and~\hypref{Decision.2}, we expect the drum sounds to be clustered together, with the majority of the synthetic noise sounds appearing away from these clusters. The projections confirmed that this is the case (see Appendix~\ref{appendix:E}). We are most interested in synthetic noise sounds which appear within or near these clusters. Do these synthetic noise sounds have drum like characteristics?\\ 
To test this, we implemented interactive projection graphs. These graphs allow interactions such as playing the samples(by hovering the cursor on-top) and movement of the scene camera. Manual inspections revealed a noticeable, positive correlation between a synthetic sound's distance and similarity of a synthetic sound to drum clusters.

\section{Pipelines and Surveys}
\label{surveys}
With our virtual synthesizer and virtual ears in place, we build two pipelines. We chose CNNLSTM-DVN for \hypref{Decision.1}, and FC-DVD,CNNLSTM-DVD and E/F for \hypref{Decision.2} to build the TPE pipeline. The MEM pipeline uses our best MEM, which simultaneously classifies sounds as drums and categorizes them. 
The TPE pipeline produces samples in the following categories: \enquote{snare}, \enquote{kick}, \enquote{hat}, \enquote{clap} and \enquote{other} (combination of rims, shakers and unusual percussive sounds). The MEM pipeline does not output the \enquote{other} category, yet the option is available to surveyors when categorizing sounds manually.   
The sounds and the corresponding synthesizer programs are saved to hard-disk. Here we assess the success rate of two different pipelines---and by extension, the virtual ears---by manual inspection of sounds categorized as drums, as well as the category of those drums. Both authors categorized a subset of the results without knowledge of the drum categorization outcomes. Additionally, each responder had the option of labeling samples as \enquote{bad} for samples that they deemed not percussive.  The percentage of sounds labeled as \enquote{bad} by the authors is a measurement of success with regards to \hypref{Decision.1}. We measure success with regards to \hypref{Decision.2}---the reliability of agreement between persons and drum categorization models--- via the Fleiss' kappa coefficient \cite{fleiss1971measuring}. The value of 0 or less for this coefficient indicates no agreement beyond random chance, and the values of 1 indicates perfect agreement. We again measure this coefficient after dropping samples that were categorized as \enquote{bad} by one or both authors. We also show measurements after dropping all \enquote{bad} and \enquote{other} samples.
\begin{table}[htbp]
 \resizebox{\linewidth}{!}{\begin{tabular}{|c c c c c c c|} 
 \hline
 Drop Rule & Size & H+H & H+FC & H+CNN & H+E/F & 3 models \\ [0.5ex] 
 \hline
 No Drops & 257 &0.37 & 0.35 & 0.36 & 0.36 & 0.28\\ 
 \hline
 Assigned \enquote{Bad} By Both & 236 & 0.31 & 0.37 & 0.37 & 0.38 & 0.30 \\
 \hline
 Assigned \enquote{Bad} By Either & 180 & 0.47 & 0.50 & 0.48 & 0.48 &  0.34 \\
 \hline
 Assigned \enquote{Bad} or \enquote{Other} By Either & 154 & 0.47 & 0.59 & 0.54 & 0.50 &  0.35 \\
 \hline
\end{tabular}}
\caption{\label{kappa_table_TPE}Table of Fleiss' kappa coefficient to measure the degree of agreement between persons (H+H), persons with FC model (H+FC), persons with CNNLSTM model, persons with all models (H+E/F), and between the 3 models. \enquote{Drop Rule} column indicates if any samples were dropped. We show the measurements after dropping samples if they are deemed bad by either or both responders. We also show measurements after dropping the \enquote{other} category along with samples deemed bad by either responder. }
\end{table}

 \begin{table}[t]
 \resizebox{\linewidth}{!}{\begin{tabular}{||c c c c||} 
 \hline
 Drop Rule & Size & H+H & H+H+MEM \\
 \hline
 No Drop & 300 & 0.34 & 0.25\\ 
 \hline
 Assigned Bad By Both & 249 & 0.20 & 0.26 \\
 \hline
 Assigned Bad By Either & 151 & 0.46 &  0.47 \\
 \hline
Assigned \enquote{Bad} or \enquote{Other} By Either  & 120 & 0.62   &  0.59 \\
 \hline
\end{tabular}}
\caption{\label{kappa_table_MEM}Table of Fleiss' kappa coefficient to measure the degree of agreement between persons (H+H) and persons with MEM (H+MEM). We also measure the agree-ability scores after dropping bad samples if both or either persons assigned the sample as such. We also measure agree-ability when all samples deemed \enquote{bad} or \enquote{other} by either person are removed.}
\end{table}

\subsection{Survey Discussion}
\label{survey2_takeaway}
Due to the discrepancy in categorization and output groups, results from the two surveys must be compared with caution. Agree-ability scores for MEM pipeline are often lower, but we reiterate that the MEM survey has two out of six (rather than one out of six in the TPE pipeline) categories which are exclusive to survey responders. Supporting this hypothesis is the notable improvement in agree-ability observed after dropping all \enquote{bad} and \enquote{other} samples. Despite the drawbacks, the results here are encouraging as the MEM pipeline remains competitive despite working with a fraction of the features to learn from and simultaneously making both decisions. Suggesting that latent representations of autoencoder networks can be used as general, low-dimensional proxies for representation of complex inputs.

\bibliographystyle{apacite}
\bibliography{CSMC_MUME_LaTeX_Template.bib}
\begin{appendices}
\chapter{Datasets}
\label{appendix:A}
\begin{table}[h!]
\centering
\begin{tabular}[width=\paperwidth]{|l|l|l|l|l|l|l|l|l|l|}
\hline
DB Name & kick & snare & clap & tom\_high & tom\_mid & tom\_low & hihat\_closed &  hihat\_open & rim \\ \hline
MixedDB & 648 & 732 & 118 & 179 & 139 &  188 & 187 & 280 & 105 \\\hline
\end{tabular}
\caption{Database 1: Mixed sources}
\label{db:self}
\end{table}

@@

%================
\chapter{Virtual Ear Classifier Definitions}
\label{appendix:C}
\section{Two Phased models}
Using the described features, we trained several neural network models for Phase 1 and 2 in the Pytorch environment. The task of Phase 1 is to separate drums from not-drums (DrumVsNotDrum, or DVN). The task of Phase 2 is to categorize drums and percussion (DrumVsDrum, or DVD). We kept our feature space small, making it viable for feature selection and model design to be done on a trial and error basis. For all models, accuracy is calculated by prediction of all test dataset labels and the loss function and optimizer are Categorical-CrossEntropy and Adam respectively. Training continues until no reduction in loss and accuracy is observed in 10 epochs.  All activation functions are PReLU:
\begin {enumerate}
\item FC-DVN: Fully connected network trained on Envelope features, reaching 97\% accuracy on our test data for Phase 1. With size of 10x5x10.
\item CNNLSTM-DVN: A combination of CNN and LSTM models, where the CNN model extracts higher level features that are fed temporally to an LSTM cell. This model is trained on spectrum data and reaches 98\% accuracy on our test set. Its structure is the combination of a CNN with 2 output channels and kernel size $(7,3)$; Followed by an LSTM model of hidden size 800 and a fully connected layer of size 20x2.
\item E/F-DVD: A fully connected model trained on a concatenation of envelope and frequency features. Reaching 80\% accuracy for 6-way drum categorization in Phase 2. Size of 50x10x2x6.
\item CNN-DVD: A CNN model trained on Spectrum features. Reaching 82\% accuracy in a 6-way drum categorization in Phase 2. A combination of a CNN model with output channel size of 4, kernel of size of 5, another CNN model with output channel size of 8 and kernel of size 3. Followed by a fully connected network of shape 100x20x6.
\item FC-DVD: Fully connected 3 layer neural net with 78\% accuracy for 6-way drum categorization in Phase 2. Size of 400x200x50.
\end{enumerate}
Unlike EE models, parameters are hand-picked and un-tuned. Higher accuracy rates in these models do not necessarily translate to higher agreeableness with surveyors. Since the positive sounds in the hearing test are sourced from a synthesizer, model accuracy on test data alone cannot be relied upon when the domain of sounds 
being categorized is switched.

@@

%==============
\chapter{AutoEncoder Structures and Hyper-Parameter Optimization}
\label{appendix:hyperparam}
% could do multi-objective pareto front stuff too
Within the context of machine learning, a model's \emph{hyper-parameters} are fixed parameters which are set before the training begins (e.g number of layers, size of layers, loss function) and are not learned during the minimization procedure~\cite{bengio2000gradient}. To assist us with the construction of our model, we defined a number of possible choices for the architecture of our model and audio-transformers and used hyper-parameter optimization to extract promising sets of values~\ref{table:hyper_params}.

@@

Using the optuna optimization tools \cite{akiba2019optuna}, we conducted 500 search trials. The trial's success is measured in their final loss value, calculated by applying the model to test data-set. Each trials consisted of 20 epoch of training using a selected set of hyper-parameters. An additional 100 trials with 40 epochs of training were conducted following the initial 500 to test the effect of longer epochs. Each trial's intermediate results (loss at every n epoch,s where $0<n<20$) were reported to a multi-armed bandit based pruner for early stoppage of unpromising trials~\cite{li2017hyperband}. We employed a tree-structured parzen estimator for better navigation of the search space~\cite{bergstra2011algorithms,akiba2019optuna} but found short reversions to a random sampling coupled with a decrease in the frequency of pruning helpful in exiting local minima. \\
\begin{table}[htbp]
